|

### 2025-07-10 output/IMPS-2025-slideshow.qmd
Drop "unidimensional items" slide
Not necessary to justify the "inverse items" generalization

:::

## Increasing monotony

::::: columns
::: column
<br>

<!-- TODO: Same color as item? -->

$$
  \begin{align}
    \mathbf{a}_1 &=
      \begin{bmatrix}
        1.5 \\
        0
      \end{bmatrix} \\
      \\
      \\
    d_1 &= 0.75
  \end{align}
$$
:::

::: column
```{r m2pl-representation-unidimensional}
#| fig-align:  center
#| fig-height: 7.2
#| fig-width:  7.2

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1.5,  0,    0.75
)

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-1, 1),
    breaks       = NULL,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = .4),
    axis.title.y.right = element_markdown(vjust = .4),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    ),
  )
```
:::
:::::

::: notes
In the first case, one may argue that we may deal with "undimensional items" or, more generally, items that have null discrimination components.

This would imply that the probability of a correct response would not change with respect to that dimension; i.e., that trait is not being measured by the item, so the isoprobability lines would be parallel to the $\theta_2$ axis (in this example.

Therefore, the probability does not "increase monotonically".

Nevertheless, with this caveat about null discrimination components, this assumption would be realistic in a "cognitive testing context": The more you know, the more likely you are to give a correct answer.
These assumptions derive from the traditional application of the model to the cognitive domain, but we want to apply multidimensional IRT to the non-cognitive domain as well, e.g. to personality measures.
So, our purpose is **to relax these two assumptions**.
:::

>>>>

These assumptions derive from the traditional application of the model to the cognitive domain, but we want to apply multidimensional IRT to the non-cognitive domain as well, e.g. to personality measures.
So, our purpose is **to relax these two assumptions**.
:::
